---
title: "ADA F22 Midterm Question 3"
author: "Daiszha Cooley"
date: "2022-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages
```{r}
pacman::p_load(tidyverse, knitr,ggplot2)
```


# Import Data Set 
```{r}
Titanic<-read_csv("https://raw.githubusercontent.com/kijohnson/Data/main/titanic.csv")
```

# Determine the Median Age by passenger class
```{r}
table(Titanic$Pclass)
# create factor version of passenger class 
Titanic$Pclass_f<-factor(Titanic$Pclass, levels = c(1,2,3), labels = c("1st class", "2nd class","3rd class"))
#confirm factor version worked 
table(Titanic$Pclass_f)
#determine median age by passenger class 
Titanic%>%
  group_by(Titanic$Pclass_f)%>%
  summarise(Median = median(Age),n = n())
```

# Plot Median, Spread of Data and Outliers (box plot) for age according to passenger class 
```{r}
AgePClassPlot<-ggplot(data = Titanic, aes(x = Age, y = Titanic$Pclass_f)) +
  geom_boxplot() +
  labs(x = "Age (yrs)", y = "Passenger Class",  title = "Age by Passenger Class")+ 
  theme_bw()
#view plot 
AgePClassPlot
```

# Report the medians by passenger class 
The median age for first class was 38.5 years old. The median age for second class was 29.0 years old. The median age for 3rd class was 24.0 years old. 

# Export plot as PNG
```{r}
ggsave(filename = "Age By Passenger Class.png", plot = AgePClassPlot,
      bg="white") 
```

# Determine the percentage of males and females by class 
```{r}
#Check class of sex -- currently character 
class(Titanic$Sex)
table(Titanic$Sex)
#create factor version of sex 
Titanic$Sex_f<-factor(Titanic$Sex,levels = c("female", "male"), labels = c("Female", "Male"))
# Check to see if it worked 
table(Titanic$Sex_f)
#Determine percentage of males and females by class 
## Create a crosstabulation of sex by class and save table 
MFbyClass<-table(Titanic$Pclass_f,Titanic$Sex_f)
## View Table 
MFbyClass
#Calculate the percentage of males and females by each class convert it to percent values and round to one decimal point 
round(100*prop.table(MFbyClass, margin = 1),1)
```

